Visual narratives of the Covid-19 pandemic

### 2022-01-17 JDSSV_manuscript_knitr.Rnw
Update graph with legend (screenshot)

While traditionally, time-series information has been limited to charts with time on the x-axis, there have been many attempts to make use of the capabilities of web graphics and animation in order to convey temporal information in less traditional time-series forms. One controversial chart appeared in the New York Times; the x-axis was wrapped around an Archimedean spiral to provide a sense of periodicity in the year-by-year evolution of case counts, as shown in \Cref{fig:nyt-spiral}. The original form of this chart is prone to the line-width illusion \citep{vanderplasSignsSineIllusion2015}, in addition to the well-known problems with polar charts \citep{hofmannGraphicalTestsPower2012,waldnerComparisonRadialLinear2020}. The re-envisioned version, which alignes counts on the outside of the spiral, and displays them as radial lines, mitigates some of the issues with the line-width illusion by directly showing each line as its own entity, facilitating direct comparisons. Still, it is more work to interpret and compare peaks on this chart than on a similar linear time series chart with the same data, as the reader must assess line width and then do a mental rotation and shift operation in order to compare with any other time period.

\begin{figure}\centering
\includegraphics[width=.45\linewidth]{nyt-spiral}\hfill\includegraphics[width=.45\linewidth]{spiral-rework}
\caption{Covid cases in the United States, 2020-2022. The original graph from the New York Times is on the left, and a reenvisioned version which is more perceptually friendly is on the right. Source: (Left) New York Times. Here's When We Expect Omicron to Peak \url{https://www.nytimes.com/2022/01/06/opinion/omicron-covid-us.html}, (Right) Covid Spiral, by Sourya Shrestha, \url{https://github.com/souryashrestha/covid_spiral}.}
\label{fig:nyt-spiral}
\end{figure}

>>>>

While traditionally, time-series information has been limited to charts with time on the x-axis, there have been many attempts to make use of the capabilities of web graphics and animation in order to convey temporal information in less traditional time-series forms. One controversial chart appeared in the New York Times; the x-axis was wrapped around an Archimedean spiral to provide a sense of periodicity in the year-by-year evolution of case counts, as shown in \Cref{fig:nyt-spiral}. The original form of this chart is prone to the line-width illusion \citep{vanderplasSignsSineIllusion2015}, in addition to the well-known problems with polar charts \citep{hofmannGraphicalTestsPower2012,waldnerComparisonRadialLinear2020}. The re-envisioned version, which aligns counts on the outside of the spiral, and displays them as radial lines, mitigates some of the issues with the line-width illusion by directly showing each line as its own entity, facilitating direct comparisons. In addition, the re-envisioned chart includes a reference line at 100K cases/day, which allows the reader to compare the severity of different waves directly. Still, it is more work to interpret and compare peaks on this chart than on a similar linear time series chart with the same data, as the reader must assess line width and then do a mental rotation and shift operation in order to compare with any other time period.

\begin{figure}\centering
\includegraphics[width=.45\linewidth]{nyt-spiral}\hfill\includegraphics[width=.45\linewidth]{spiral-rework}
\caption{Covid cases in the United States, 2020-2022. The original graph from the New York Times is on the left, and a reenvisioned version which is more perceptually friendly is on the right. Source: (Left) New York Times \url{https://www.nytimes.com/2022/01/06/opinion/omicron-covid-us.html}, (Right) Covid Spiral, by Sourya Shrestha, \url{https://github.com/souryashrestha/covid_spiral}.}
\label{fig:nyt-spiral}
\end{figure}
